RBE/CS549 Project: Boat Tracking/Detection

### 2017-12-13 paper/main.tex
Fix spelling errors, make some lines 80 characters

\subsection{Naiive K-Means Classifier}
We initially started with a basic two-class k-means classifier in order to observe how well it would perform.  The classifier was trained on the first image in the video sequence, and the l2 distance to the two cluster centers was used to classify each pixel in subsequent frames.  Surprisingly enough, the 2-class k-means classifier worked relatively robustly on the videos that we provided the algorithm, as show in figure \ref{fig:kmeans}.  The water pixels are mostly labeled as one class, and pixels that not water are mostly labeled as another class with a little bit of misclassification noise.  The boat and it's wake are clearly labeled as potential obstacles to track, which is what we want. The land areas are labeled as water pixels, but for our implementation we want to ignore tracking the land anyways. There is a slight bit of noise in the water reflections towards the right side of figure \ref{fig:kmeans}, but this naiive method still provides very useful information. \\
\newline
Kmeans does not always associate the same class identifier with the water and non-water pixels each time, so we must perform an intermediary step to detect which class identifier it associates with water pixels.  To do so, we find the number of pixels in each class in the bottom half of the image, and use the sums to vote for which class is associated with water pixels. This assumes that the boat is upright, and that a majority of the pixels in the bottom half of the image are water, and holds true for all of the cases we observed.  If the water class ID is not what was expected, the cluster centers are swapped so that the water class ID is always a known value.

\begin{figure}
\label{fig:kmeans}
\includegraphics[width=7.8cm]{hsv_kmeans2_orig}
\includegraphics[width=8.5cm]{hsv_kmeans2_result}
\centering
\caption{Naiive K-Means Classifier Results}
\end{figure}

@@

\subsection{Sky Suppression}
The k-means classifier provides information about which pixels are water and which pixels are not water, but we would also like the algorithm to only focus on potential obstacles in the water.  The region of the sky shown in figure \ref{fig:kmeans} is classified as a non-water pixel, and we would like to suppress this response to avoid detecting the sky as an obstacle. With only the wave reflection suppression confidences, most of the sky is suppressed, but tends to have higher confidences near land as shown in figure \ref{fig:skysuppress}.  Since the sky pixels are all connected, and define a very large region, we used the Matlab bwareaopen function that removes binary pixels in connected areas that are smaller than a particular size.  By looking only at pixels that are classfied as "not water", we can detect the sky region by filtering for pixel clusters that are larger than 1/4 of the area of the image, and set the confidence values for those pixels to zero to suppress them. In some cases, potential objects to track were touching the sky region and the bwareaopen results tended to remove those object to track.  To fix this, an erode operation is done on the non-water class pixels to better separate the sky and potential objects to track.

\begin{figure}
\label{fig:skysuppress}
\includegraphics[height=5cm]{hsv_confidence}
\includegraphics[height=5cm]{hsv_kmeans2_suppressed}
\centering
\caption{Sky Supression Before and After}
\end{figure}

\subsection{Confidence Image Results}
When the above confidence values and suppressions described above are combined, a final confidence value image is produced. Pixels that are more likely to be objects to track have higher confidence values that pixels that are more likely to be water areas. Figure \ref{fig:boatconf} shows the detector correctly detecting a boat in the image.  Figure \ref{fig:towerconf} shows the detector correctly detecting a tower in the ditsance.  Figure \ref{fig:buoyconf} shows the detector identifying a buoy off in the distance.

@@

\section{Future Results}

>>>>

\subsection{Naive K-Means Classifier}
We initially started with a basic two-class k-means classifier in order to
observe how well it would perform.  The classifier was trained on the first
image in the video sequence, and the l2 distance to the two cluster centers was
used to classify each pixel in subsequent frames.  Surprisingly enough, the
2-class k-means classifier worked relatively robustly on the videos that we
provided the algorithm, as show in figure \ref{fig:kmeans}.  The water pixels
are mostly labeled as one class, and pixels that not water are mostly labeled as
another class with a little bit of misclassification noise.  The boat and it's
wake are clearly labeled as potential obstacles to track, which is what we want.
The land areas are labeled as water pixels, but for our implementation we want
to ignore tracking the land anyways. There is a slight bit of noise in the water
reflections towards the right side of figure \ref{fig:kmeans}, but this naive
method still provides very useful information.

Kmeans does not always associate the same class identifier with the water and
non-water pixels each time, so we must perform an intermediary step to detect
which class identifier it associates with water pixels.  To do so, we find the
number of pixels in each class in the bottom half of the image, and use the sums
to vote for which class is associated with water pixels. This assumes that the
boat is upright, and that a majority of the pixels in the bottom half of the
image are water, and holds true for all of the cases we observed.  If the water
class ID is not what was expected, the cluster centers are swapped so that the
water class ID is always a known value.

\begin{figure}
\label{fig:kmeans}
\includegraphics[width=7.8cm]{hsv_kmeans2_orig}
\includegraphics[width=8.5cm]{hsv_kmeans2_result}
\centering
\caption{Naive K-Means Classifier Results}
\end{figure}

@@

\subsection{Sky Suppression}
The k-means classifier provides information about which pixels are water and
which pixels are not water, but we would also like the algorithm to only focus
on potential obstacles in the water.  The region of the sky shown in figure
\ref{fig:kmeans} is classified as a non-water pixel, and we would like to
suppress this response to avoid detecting the sky as an obstacle. With only the
wave reflection suppression confidences, most of the sky is suppressed, but
tends to have higher confidences near land as shown in figure
\ref{fig:skysuppress}.  Since the sky pixels are all connected, and define a
very large region, we used the Matlab bwareaopen function that removes binary
pixels in connected areas that are smaller than a particular size.  By looking
only at pixels that are classified as "not water", we can detect the sky region
by filtering for pixel clusters that are larger than 1/4 of the area of the
image, and set the confidence values for those pixels to zero to suppress them.
In some cases, potential objects to track were touching the sky region and the
bwareaopen results tended to remove those object to track.  To fix this, an
erode operation is done on the non-water class pixels to better separate the sky
and potential objects to track.

\begin{figure}
\label{fig:skysuppress}
\includegraphics[height=5cm]{hsv_confidence}
\includegraphics[height=5cm]{hsv_kmeans2_suppressed}
\centering
\caption{Sky Suppression Before and After}
\end{figure}

\subsection{Confidence Image Results}
When the above confidence values and suppressions described above are combined,
a final confidence value image is produced. Pixels that are more likely to be
objects to track have higher confidence values that pixels that are more likely
to be water areas. Figure \ref{fig:boatconf} shows the detector correctly
detecting a boat in the image.  Figure \ref{fig:towerconf} shows the detector
correctly detecting a tower in the distance.  Figure \ref{fig:buoyconf} shows
the detector identifying a buoy off in the distance.

@@

\section{Future Work}
